Explorations to Optimize the Parareal Algorithm \\to Solve ODEs in Given Applications

### 2014-12-10 paper.tex
minor changes to paper
added more stuff to paper about error

%ssssFile: paper.tex
%ss Created: Sat Dec 06 04:00 PM 2014 E
%
\documentclass[letterpaper,12pt]{article}
% Change the header if you're going to change this.

@@

\tableofcontents

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
a generalalized algorithm which allows for parallelization in time. It does so by using a
cheaper ($g_{\Delta t}$)---lower order or lesser resolution, approximation first which it then corrects by evaluating segments with a finer method ($g_{\textnormal{fine}}$). The correction step from the approximation can be done in parallel. Multiple iterations can be performed to achieve an error provably asymptotically equal to a full serial computation of $g_{\textnormal{fine}}$. The exact number of iterations required to achieve certain levels of convergence depends on the problem and must be tuned top optimize for speedup for a given system. In summary, the algorithm is $k$ repetitions of a
correction process which requires running updated by a finer method run in parallel for subsections.

\begin{algorithm}[t]
    \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finersscheme $g_{\textnormal{fine}}$}
    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
    Compute the corrections $\delta g_n(u^1_n) = g_{\textnormal{fine}}(t_n,
    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
    u^2_n) + \delta g_n(u^1_n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
    = u^1_0$ as the initial condition\;
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

@@

@@

TODO BKSIM? ASHI carry?

TODO BKSIM

Assume the coarse operator $g_{\Delta t}$ is order $m$ and is Lipshitz, and the
fine solution operator $g_{\textnormal{fine}}$ is a sufficiently accurate
approximation to the analytic operator so we may replace $g_{\textnormal{fine}}
\to g$.

\emph{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. Can be proved by
induction.

SHALL WE DO IT THEN?

FROM THE SLIDES? + Adjust?
TODO BKSIM? ASHI carry?

Show that this parareal method will approach, with large enough $k$ (enough
iterations) to approach the error of the fine method. However, with too large of
a $k$ and a lower quality factor (ratio of fine to coarse), the time for the
parareal could potentially take longer than the direct serial computation.

@@

\begin{equation}
t + k(t + \frac{Qt}{N}).
\end{equation}

@@

We wished to explore certain tests that would span the space of possible
parameters to the parareal as well as possible applications in different ODE
systems. Because there were so many different versions and parameteres for the parareal to test,

@@

\begin{figure}
\includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
\caption{Running time vs. number of CPUs.}
\label{fig:run_v_cpu}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
\caption{Error vs. number of correction steps.}
\label{fig:err_v_k}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
\caption{Error vs. quality factor of the fine operator}
\label{fig:err_v_q}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
\caption{Running time as a function of the number of correction steps.}
\label{fig:run_v_k}
\end{figure}

@@

>>>>

%File: paper.tex
%Created: Sat Dec 06 04:00 PM 2014 E
%
\documentclass[letterpaper,12pt]{article}
% Change the header if you're going to change this.

@@

%\tableofcontents

%\clearpage

@@

The parareal algorithm, developed by Lions, Maday and Turinici in 2001, CITE is
a generalized algorithm which allows for parallelization in time. It does so
by using a cheaper ($g_{\Delta t}$)---a lower order or lesser resolution
approximation first, which it then corrects by evaluating segments with a finer
method $g_{\textnormal{fine}}$. The correction step from the approximation can
be done in parallel. Multiple iterations can be performed to achieve an error
provably asymptotically equal to a full serial computation of
$g_{\textnormal{fine}}$. The exact number of iterations required to achieve
certain levels of convergence depends on the problem and must be tuned top
optimize for speedup for a given system. In summary, the algorithm is $k$
repetitions of a correction process which requires running updated by a finer
method run in parallel for subsections.

\begin{algorithm}[t]
    \KwIn{Temporal discretization $t_n = t_0 + n \Delta t, \, n = 1,2,\ldots,N$}
    \KwIn{Coarse scheme $g_{\Delta t}$}
    \KwIn{Finer scheme $g_{\textnormal{fine}}$}
    Compute $u^1_{n+1} = g_{\Delta t}(t_n, u^1_n)$\;
    Compute the corrections $\delta g(u^1_n) = g_{\textnormal{fine}}(t_n,
    u^1_n) - g_{\Delta t}(t_n, u^1_n)$ in parallel\;
    Add the prediction and correction terms as $u^2_{n+1} = g_{\Delta t}(t_n,
    u^2_n) + \delta g(u^1_n)$\;
    Repeat steps 2 and 3, incrementing the iteration label and using $u^{k+1}_0
    = u^1_0$ as the initial condition\;
 \caption{Parareal}
 \label{alg:parareal}
\end{algorithm}

@@

Assume the coarse operator $g_{\Delta t}$ is Lipschitz and order $m$: for some
constant $L$, 
\[ |g_{\Delta t}(t_n, u) - g_{\Delta t}(t_n, v)| \leq (1 + L \Delta t) |u - v|
\, \forall t \in (0, t_n), \]
\[ |u(t_n) - u^1_N | \leq C(\Delta t)^m |u_0|. \]

Additionally, assume the function $u$ is bounded on $(0, t_n)$. Furthermore,
assume the fine solution operator $g_{\textnormal{fine}}$ is a sufficiently
accurate approximation to the analytic operator so we may replace
$g_{\textnormal{fine}} \to g$.

\textbf{Theorem}: The order of accuracy of the Parareal method with coarse
solution operator $g_{\Delta t}$ and fine operator $g$ is $mk$. We can prove
this using induction [Bal].

\emph{Proof}: By induction.

Case $k = 1$: This is just the coarse operator, which is order $m$.

Assume that this is true for $k$, that $|u(t_N) - u^k_N| \leq C(\Delta t)^{mk}
|u_0|$.

For the case $k + 1$, we have
\[
\begin{aligned}
    |u(t_N) - u^{k+1}_N| &= 
    |g(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1}) - \delta g(u^k_{N-1})| \\
    &= |g_{\Delta t}(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1}) - \delta
    g(u^k_{N-1}) + \delta g(u(t_{N-1}))| \\
    &\leq |g_{\Delta t}(u(t_{N-1})) - g_{\Delta t}(u^{k+1}_{N-1})| - |\delta
    g(u^k_{N-1}) + \delta g(u(t_{N-1}))| \\
    &\leq (1 + C \Delta t) |u(t_{N-1} - u_{N-1}^{k+1}| + C(\Delta t)^{m+1} |
    u_{N-1}^k - u(t_{N-1}) | \\
    &\leq (1 + C \Delta t) |u(t_{N-1}) - u_{N-1}^{k+1}| + C(\Delta t)^{m(k+1)+1}
    | u_0 | \\
\end{aligned}
\]

We can continue to expand $|u(t_{N-1} - u_{N-1}^{k+1}|$ and get time indices of
$N-1, N-2, \ldots, 2, 1$. This implies that 
\[ |u(t_N) - u^{k+1}_N| \leq C(\Delta t)^{m(k+1)} |u_0|, \]
as desired.

With the assumptions of the previous section, this parareal method will
approach, with large enough $k$ (correction iterations) to approach the error of
the fine method. However, there is a time vs. accuracy tradeoff. Let $Q$ be the
\emph{Quality Factor} for $g_{\Delta t}$ and $g_{\textnormal{fine}}$. Suppose
that, for constant number of processors and $\Delta t$, $g_{\Delta t}$ runs in
time $T$. Then, $g_{\textnormal{fine}}$ runs in time $QT$. With too large of a
$k$ and a lower $Q$, the time for the Parareal could potentially take longer
than the direct serial computation. We note that there could be other
relationships between the time of the coarse and fine schemes other than
multiplication by a constant term.

Disregarding the time relationship between $g_{\Delta t}$ and
$g_{\textnormal{fine}}$, we can look at what happens to the error as $k \to N$
(recall that $N$ is the total number of time points in the time discretization).

For $k = 1, 2, \ldots$ we have 
\[u_{n+1}^{k+1} = g_{\Delta t} (t_n, u_n^{k+1}) + \big(
g_{\textnormal{fine}}(t_n, u_n^k) - g_{\Delta t}(t_n, u_n^k). \]

As $k \to N$, the Parareal algorithm gives $u_{n}^{k+1} = u^k_n$, so the order
of accuracy approaches that of $g_{\textnormal{fine}}$.

% I don't understand the above part ~ Andy

@@

ASHI carry?

@@

\begin{equation}
t + k \left(t + \frac{Qt}{N} \right).
\end{equation}

@@

We wished to explore certain tests that would span the space of possible
parameters to the parareal as well as possible applications in different ODE
systems. Because there were so many different versions and parameters for the Parareal to test,

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.75\textwidth]{data/runtime_vs_cpus.pdf}
\caption{Running time vs. number of CPUs.}
\label{fig:run_v_cpu}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_corrections.pdf}
\caption{Error vs. number of correction steps.}
\label{fig:err_v_k}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.75\linewidth]{data/error_vs_qualityfactor.pdf}
\caption{Error vs. quality factor of the fine operator}
\label{fig:err_v_q}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.75\linewidth]{data/runtime_vs_corrections.pdf}
\caption{Running time as a function of the number of correction steps.}
\label{fig:run_v_k}
\end{center}
\end{figure}

@@

%\bibliographystyle{unsrt}
%\bibliography{poster/sample}

Guilliame Bal paper

Stability of the Parareal Algorithm by Staff et al
